Coherent optical creation of a single molecule

### 2020-12-07 paper.tex
Update on Overleaf.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a.
The scheme does not rely on a Feshbach resonance,
molecular states bound by a few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman scheme,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states to balance the two Rabi frequencies as much as possible.
This system-independent approach creates molecules atom-by-atom with full quantum state control.

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_a$ and $\Omega_m$ with one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi Rate is given by $\Omega_a\Omega_m / (2\Delta)$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced ($\Omega_a/\Omega_m \ll 1$) %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and scattering losses are dominated by the final state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning $ \Delta $ of $80$ to $200$~GHz, causing the target molecular state to scatter both beams nearly equally with a total rate $ \Gamma_e \Omega_m^2 / (2\Delta^2)$, where $ \Gamma_e $ is the excited-state linewidth~\footnote{The two beams have equal power to maximize the Raman Rabi rate at fixed total power, resulting in a factor of 2 for scattering two beams.}.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to Stark shift ($\Omega_a/\Omega_m$)~\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}. Notably, both figures of merit improve with a larger ratio $\Omega_a/\Omega_m$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred \TR{Mention 2-photon here?}. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

%% Transfer scheme
% After the atoms are prepared in the $\ket{\uparrow_{\Na}\uparrow_{\Cs}} $ hyperfine combination, we then perform the Raman transfer.
To transfer the atom pair into the target weakly bound molecular state,
we modulate the tweezer beam with a second frequency near $770$~MHz, as shown in Fig.~\ref{f-beampath}.
The dual use of the tweezer beam for confinement and Raman transfer not only minimizes photon scattering,
but also allows a tight focus to minimize the transfer duration. A tweezer frequency far detuned (by $-151~\mathrm{GHz}$) from the 3 observed rotational~\TR{spectral?} lines at $v' = 0 $ reduces resonant scattering~\cite{Liu2019}.
%Furthermore, two filters, each with a linewidth (FWHM) of $50~\mathrm{GHz}$,
%clean the laser spectrum and prevent broadband noise from causing unwanted excitation.
%As shown in Fig.~\ref{f-beampath}, one filter immediately follows the laser, while the second filter precedes the microscope objective for final cleanup of the laser spectrum.
%The filters reduce photon scattering rate by a factor of $2$ (\textcolor{red}{KW: Since we mention this again later when we explore the causes of our lower than expected transfer efficiency, do we need to mention this reduction here? or the possible causes?}). While we have not fully characterized the sources
%of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities.
% After the total tweezer power is set to the desired value,
% we smoothly ramp down the power of one frequency in the tweezer
% while simultaneously ramping up the power of the other frequency
% so that the total tweezer power remains unchanged.
% Both frequencies are kept on for a specified duration before the process is reversed
% and the tweezer returns to a single frequency.~\TR{???}

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-lifetime.pdf}
  \caption{
    (a) Direct measurement of molecule lifetime with about $3.75~\mathrm{mW}$ optical power.
    Molecule survival is detected by dissociating back to atoms via a second Raman transition.
    The lifetime is consistent with the $0.199(9)~\mathrm{ms}$
    measured from the Raman transition data.
    Inset: pulse sequence for the lifetime measurement.
    (b) Two-body atom lifetime of $5(1)~\mathrm{ms}$
    in a trap of depth $3.75~\mathrm{mW}$ caused by off-resonance photoassociation.
    This is used to improve the fitting of the Raman transfer data.
    Inset: Atomic scattering rate scales as
    $P_\textrm{tweezer}^{2.58}\times\!2\pi\!\times29.3(17)~\mathrm{mHz/mW^{2.58}}$ on a log-log scale;
    this is consistent with a two-photon scattering process.
    We have not measured a clear dependency of the loss rate on the tweezer detuning.
    \label{f-lifetime}}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|c|c|c|c|}
    $P~(\mathrm{mW})$&$0.75$&$1.5$&$3.75$\\\hline
    $f_{\mathrm{PA}0}~(\mathrm{GHz})$&\multicolumn{3}{|c|}{$288711.8$}\\\hline
    $a~(\mathrm{2\pi\times MHz})$&$770.20452(6)$&$770.2081(1)$&$770.1943(3)$\\
    $b~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$-12.46(2)$&$-24.44(3)$&$-60.66(8)$\\\hline
    $c~(\mathrm{2\pi\times kHz})$&$0.29(2)$&$0.63(4)$&$2.4(2)$\\
    $d~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$0.115(4)$&$0.275(6)$&$0.95(3)$\\ \hline
    $\Omega_R~(\mathrm{2\pi\times kHz})$ & & & X.YY \\
    $\Omega_m~(\mathrm{2\pi\times MHz})$ & & & 348.3(3) \\
    $\Omega_a~(\mathrm{2\pi\times kHz})$ & & & 5.5(2)
  \end{tabular}
  \caption{Fitting results for Fig.~\ref{f-det}(a,b). $\Omega_R$ is reported at $-151~\mathrm{GHz}$ detuning from the $v' = 0$ state.  The coefficients $\Omega_m$ and $\Omega_a$ are in broad agreement with our calculation and their ratio $63$ is near the theory prediction of $78$.  The measured Rabi rate $\Omega_R$ is ZZ~\% of $\Omega_m \Omega_a/(2\Delta)$ due to interference from further-detuned Raman processes.
    \label{tab:f-det:fit}}
\end{table}

@@

The measured single-photon Rabi frequencies of Table~I are in broad agreement with calculations (see Supplementary Materials). Known single-photon scattering rates of Na and Cs atoms together with an independent measurement (via photoassociation spectroscopy) of the $v'=0$ excited state natural linewidth of no more than $50~\mathrm{MHz}$ indicate that single-photon scattering from the molecular or atomic state do not explain the decoherence in Fig. 3b~\TR{long sentence}.  Below, we further investigate the loss of coherence.

@@

Our technique can be applied to form a diverse set of molecular species,
because it does not rely on a magnetic Feshbach resonance, states bound by only a few MHz,
or a narrow excited state. The formation of weakly bound molecules is a key step
in forming rovibrational ground state molecules. By scaling up to many optical tweezers~\cite{Endres2016, Barredo2018,  PhysRevLett.122.203601}, large arrays with arbitrary geometry of highly controlled molecules can be achieved. These highly controlled molecules comprise a flexible platform for quantum simulation and quantum computing applications. \textcolor{blue}{(End with a sentence on molecules for future quantum applications?)}
% Combined with real-time rearrangement~\cite{Barredo2016, Endres2016},
% defect-free arrays of highly controlled molecules comprise a promising
% and flexible platform for quantum simulation and quantum computing applications.

@@

>>>>

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a.
The scheme does not rely on a Feshbach resonance,
molecular states bound by a few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman scheme,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states to balance the two Rabi frequencies as much as possible.
This approach applies to a variety of molecules that can be created atom-by-atom with full quantum state control.

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_a$ and $\Omega_m$ with one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi Rate is given by $\Omega_a\Omega_m / (2\Delta)$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced ($\Omega_a/\Omega_m \ll 1$) %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and scattering losses are dominated by the final state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning $ \Delta $ of $80$ to $200$~GHz, causing the target molecular state to scatter both beams nearly equally with a total rate $ \Gamma_e \Omega_m^2 / (2\Delta^2)$, where $ \Gamma_e $ is the excited-state linewidth~\footnote{The two beams have equal power to maximize the Raman Rabi rate at fixed total power, resulting in a factor of 2 for scattering two beams.}.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to light shift ($\Omega_a/\Omega_m$)~\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}. Notably, both figures of merit improve with a larger ratio $\Omega_a/\Omega_m$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
%However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

%% Transfer scheme
% After the atoms are prepared in the $\ket{\uparrow_{\Na}\uparrow_{\Cs}} $ hyperfine combination, we then perform the Raman transfer.
To transfer the atom pair into the target weakly bound molecular state,
we modulate the tweezer beam with a second frequency near $770$~MHz, as shown in Fig.~\ref{f-beampath}.
The dual use of the tweezer beam for confinement and Raman transfer not only minimizes photon scattering,
but also allows a tight focus to minimize the transfer duration. A tweezer frequency far detuned (by $-151~\mathrm{GHz}$) from $v' = 0 $ ($<50$~MHz natural linewidth) reduces resonant scattering~\cite{Liu2019}.
%Furthermore, two filters, each with a linewidth (FWHM) of $50~\mathrm{GHz}$,
%clean the laser spectrum and prevent broadband noise from causing unwanted excitation.
%As shown in Fig.~\ref{f-beampath}, one filter immediately follows the laser, while the second filter precedes the microscope objective for final cleanup of the laser spectrum.
%The filters reduce photon scattering rate by a factor of $2$ (\textcolor{red}{KW: Since we mention this again later when we explore the causes of our lower than expected transfer efficiency, do we need to mention this reduction here? or the possible causes?}). While we have not fully characterized the sources
%of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities.
% After the total tweezer power is set to the desired value,
% we smoothly ramp down the power of one frequency in the tweezer
% while simultaneously ramping up the power of the other frequency
% so that the total tweezer power remains unchanged.
% Both frequencies are kept on for a specified duration before the process is reversed
% and the tweezer returns to a single frequency.~\TR{???}

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-lifetime.pdf}
  \caption{
    (a) Direct measurement of molecule lifetime with about $X.XX~\mathrm{mW}$ optical power.
    Molecule survival is detected by dissociating back to atoms via a second Raman transition.
    The lifetime is consistent with the $0.199(9)~\mathrm{ms}$
    measured from the Raman transition data.
    Inset: pulse sequence for the lifetime measurement.
    (b) Two-body atom lifetime of $5(1)~\mathrm{ms}$
    in a trap of depth $X.XX~\mathrm{mW}$ caused by off-resonance photoassociation.
    This is used to improve the fitting of the Raman transfer data.
    Inset: Atomic scattering rate scales as
    $P_\textrm{tweezer}^{2.58}\times\!2\pi\!\times29.3(17)~\mathrm{mHz/mW^{2.58}}$ on a log-log scale;
    this is consistent with a two-photon scattering process.
    We have not measured a clear dependency of the loss rate on the tweezer detuning.
    \label{f-lifetime}}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|c|c|c|c|}
    $P~(\mathrm{mW})$&$0.75$&$1.5$&$3.75$\\\hline
    $f_{\mathrm{PA}0}~(\mathrm{GHz})$&\multicolumn{3}{|c|}{$288711.8$}\\\hline
    $a~(\mathrm{2\pi\times MHz})$&$770.20452(6)$&$770.2081(1)$&$770.1943(3)$\\
    $b~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$-12.46(2)$&$-24.44(3)$&$-60.66(8)$\\\hline
    $c~(\mathrm{2\pi\times kHz})$&$0.29(2)$&$0.63(4)$&$2.4(2)$\\
    $d~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$0.115(4)$&$0.275(6)$&$0.95(3)$\\ \hline
    $\Omega_R~(\mathrm{2\pi\times kHz})$ & 0.49(2) & 1.18(3) & 3.97(9) \\
    $\Omega_m~(\mathrm{2\pi\times MHz})$ & & & 348.3(3) \\
    $\Omega_a~(\mathrm{2\pi\times kHz})$ & & & 5.5(2)
  \end{tabular}
  \caption{Fitting results for Fig.~\ref{f-det}(a,b). $\Omega_R$ is reported at $-151~\mathrm{GHz}$ detuning from the $v' = 0$ state.  The coefficients $\Omega_m$ and $\Omega_a$ are in broad agreement with our calculation and their ratio at $3.75~\mathrm{mW}$ is $0.016$, which is near the theory prediction of $0.013$.  The measured Rabi rate $\Omega_R$ is ZZ~\% of $\Omega_m \Omega_a/(2\Delta)$ due to interference from further-detuned Raman processes.
    \label{tab:f-det:fit}}
\end{table}

@@

The measured single-photon Rabi frequencies of Table~I are in broad agreement with calculations (see Supplementary Materials). From the full calculation of the expected scattering rate including all excited states
%Known single-photon scattering rates of Na and Cs atoms together with an independent measurement (via photoassociation spectroscopy) of the $v'=0$ excited state natural linewidth of no more than $50~\mathrm{MHz}$ indicate that single-photon scattering from the molecular or atomic state do not explain the decoherence in Fig. 3b~\TR{long sentence}.  Below, we further investigate the loss of coherence.

@@

Our technique can be applied to form a diverse set of molecular species,
because it does not rely on a magnetic Feshbach resonance, states bound by only a few MHz,
or a narrow excited state. The formation of weakly bound molecules is a key step
in forming rovibrational ground state molecules. By scaling up to many optical tweezers~\cite{Endres2016, Barredo2018,  PhysRevLett.122.203601}, large arrays with arbitrary geometry of highly controlled molecules can be achieved. These molecules comprise a flexible platform for quantum simulation and quantum computing applications. \textcolor{blue}{(End with a sentence on molecules for future quantum applications?)}
% Combined with real-time rearrangement~\cite{Barredo2016, Endres2016},
% defect-free arrays of highly controlled molecules comprise a promising
% and flexible platform for quantum simulation and quantum computing applications.

@@

\TR{Broader statement on experimentally possible vs. theoretical detunings.}
